SoK: The Evolution of Defenses Against Web Application Attacks

### 2014-11-04 docs/questioning.tex
changes to Section III

Web application attacks may involve security misconfigurations,
broken authentication and session management, or other issues.
Some of the most dangerous web application attacks, though,
are associated with the processing of external input data
that may contain malicious code, a vulnerability known as
code injection.

Such attacks include Cross-Site Scripting ({\sc xss})
attacks~\cite{SG07}, {\sc sql} injection attacks~\cite{RL12b},
Cross-Site Request Forgery ({\sc csrf}) attacks~\cite{LZRL09}, and
others. Attacks like these have been topping the vulnerability lists
of numerous bulletin providers like the Open Web Application Security
Project\footnote{\url{https://www.owasp.org/index.php/Category:OWASP_Top_Ten_Project}}
({\sc owasp}) and
{\sc mitre}\footnote{\url{http://cwe.mitre.org/top25/}} for several
years. Consider the case of {\sc owasp}'s Top Ten project whose main
goal is to raise awareness about web application security by
identifying some of the most critical risks facing organizations which
is referenced by numerous researchers. In its three consecutive Top
Ten lists (2007, 2010, 2013), different executable source code-driven
injection attacks dominate the top five positions. Malicious users
find new ways to bypass defense mechanisms by using a variety of
techniques despite the numerous countermeasures that are being
introduced. Note that the number of systems developed to counter
{\sc sql} injection attacks until 2006 was more than twenty~\cite{HVO06}.
Since then the number has doubled.
Researchers have also indicated that the number of such attacks
is steadily increasing in recent years~\cite{SSL12}.

In this paper we initially show how the above attacks can be seen from
a {\it common perspective}. We achieve this by providing
an exploitation model that highlights the following fact:
the majority of the steps needed to perform application attacks
based on executable source code injection are
common.\footnote{Note that we leave out attacks that are based on
binary code injection (i.e. buffer
overflows --- see Section~\ref{sec:attacks}).}
This is validated by the fact that there are
protection mechanisms that can detect more than
one of these attacks. Then, we {\it categorize}
the top 34 protection mechanisms that detect such attacks and
are presented in publications cited more than twenty times.
For the sake of our study, we choose
to make some exceptions and include some of the
latest mechanisms that were presented in top security
conferences even if the corresponding publications have
been cited less than twenty times. Finally, we {\it analyze}
each mechanism based on the following dimensions:
\begin{itemize}
\item {\it Diagnostic Performance} As long
as we examine protection mechanisms that detect
attacks, the non-existence of false positive and
negative alarms is a reasonable requirement.
\item {\it Computational Performance} We examine
the overhead of the mechanisms
that affect the experience of a web user who actually
uses the protected application.
\item {\it Ease of use} We examine if the protection
mechanism is practical in terms of deployment
and can be easily adopted by security experts.
\item {\it Security} How resilient is the system to
attackers that know how it works and wants to circumvent it.
\item {\it Detection Locus}
\item {\it Availability}
\end{itemize}

\noindent
All the aforementioned requirements are considered critical
when building security mechanisms that protect
applications~\cite{A01,A00}. Based on this analysis
we can identify the advantages and disadvantages of the
various mechanisms and enumerate some of their common
characteristics. We can also draw useful conclusions
about the various categories and see how they compare
to each other. A worrying observation is that sometimes,
researchers test the diagnostic performance of their
mechanisms in a poor manner. This turns out as very
critical because some of these mechanisms can actually
be bypassed. In addition, researchers do
not make their mechanisms available to the community,
thus no one can either use their mechanism,
or replicate their experiments.

@@

\begin{enumerate}
\item We provide an exploitation model that puts
application attacks based on executable source code injection
under a common perspective.
\item We categorize and analyze previously proposed
defenses based on criteria that are considered critical
when building protection mechanisms.
\item We discuss some serious issues raised from our
findings. Such can be the main reasons why
the mechanisms are not adopted in practice.
\item We discuss what are the necessary criteria
for new mechanisms and how our work can help
researchers to design new countermeasures.
\end{enumerate}

@@

Bratus et al.~\cite{BLSPS11} portray the issue in a generic fashion:
{\it ``unexpected (and unexpectedly powerful) computational models
inside targeted systems, which turn a part of the target into a
so-called `weird machine' programmable by the attacker via crafted inputs
(a.k.a. `exploits')''}.
An example of the above definition is the
following: The code fragment below defines the operation of
addition in the Scheme programming language~\cite{AS96,D09}:

\begin{verbatim}
(define (add x y) (+ x y))
\end{verbatim}

\noindent
Consider the case where instead of a number, a function that leads to an
endless loop is passed as an argument by the user. This will cause
the interpreter to enter
an endless loop and lead to a denial of service.
% Keep in mind that a {\sc cia} does not require a programming
% language with first-class functions.
The intuition here is that {\it ``every application that copies
untrusted input verbatim into an output
program is vulnerable to code injection''}.
Ray and Ligatti~\cite{RL12b} actually proved the above
claim based on formal language theory.

@@

{\sc xss} attacks fall ino two categories, {\it non-persistent} and
{\it persistent}. Non-persistent {\sc xss} attacks take place when the
data provided by a user is used immediately by server-side scripts to
parse and display a page of results, without properly sanitizing the
request ({\sc step}: {\sc p-xss} 8 \text{\textbar} {\sc n-xss} 3).
Thus, in a non-persistent {\sc xss} attack the injected code
never reaches the database and immediately becomes a part of the
content that is sent back to a web user.

>>>>

Web application attacks may involve security misconfigurations, broken
authentication and session management, or other issues. Some of the
most dangerous web application attacks are associated with the
processing of external input data that can contain malicious code, a
vulnerability known as code injection.

Such attacks include Cross-Site Scripting ({\sc xss})
attacks~\cite{SG07}, {\sc sql} injection attacks~\cite{RL12b},
Cross-Site Request Forgery ({\sc csrf}) attacks~\cite{LZRL09}, and
others. Attacks like these have been topping the vulnerability lists
of numerous bulletin providers like the Open Web Application Security
Project\footnote{\url{https://www.owasp.org/index.php/Category:OWASP_Top_Ten_Project}}
({\sc owasp}) and {\sc
  mitre}\footnote{\url{http://cwe.mitre.org/top25/}} for several
years. Consider the case of {\sc owasp}'s Top Ten project whose main
goal is to raise awareness about web application security by
identifying some of the most critical risks facing organizations. The
Top Ten project is popular, referenced by numerous researchers. In its
three consecutive Top Ten lists (2007, 2010, 2013), different
executable source code-driven injection attacks dominate the top five
positions. Malicious users find new ways to bypass defense mechanisms
by using a variety of techniques despite the numerous countermeasures
that are being introduced. Note that the number of systems developed
to counter {\sc sql} injection attacks until 2006 was more than
twenty~\cite{HVO06}. Since then the number has doubled. Researchers
have also indicated that the number of such attacks is steadily
increasing in recent years~\cite{SSL12}.

In this paper we initially show how the above attacks can be seen from
a {\it common perspective}. We achieve this by providing an
exploitation model that highlights the following fact: the majority of
the steps needed to perform application attacks based on executable
source code injection are common.\footnote{Note that we leave out
  attacks that are based on binary code injection (i.e. buffer
  overflows --- see Section~\ref{sec:attacks}).} This is validated by
the fact that there are protection mechanisms that can detect more
than one of these attacks. Then, we {\it categorize} a selection of
these protection mechanisms. The selection comprises protection
mechanisms that detect such attacks and are presented in publications
cited more than twenty times, and some of the latest mechanisms that
have been presented in top security conferences that have been cited
less than twenty times. We included the second group so that we would
not penalize recent research. Finally, we {\it analyze} each mechanism based
on the following dimensions:
\begin{itemize}
\item {\it Diagnostic Performance} Protection mechanisms are as good
  as their detection capability; this requires low false positive and
  false negative rates.
\item {\it Computational Performance} We examine
the overhead of the mechanisms
that affect the experience of a web user who actually
uses the protected application.
\item {\it Ease of use} We examine if the protection
mechanism is practical in terms of deployment
and can be easily adopted by security experts.
\item {\it Security} How resilient is the system to
attackers that know how it works and wants to circumvent it.
\item {\it Detection Locus}
\item {\it Availability}
\end{itemize}

\noindent
All the aforementioned requirements are considered critical
when building security mechanisms that protect
applications~\cite{A01,A00}. Based on this analysis
we can identify the advantages and disadvantages of the
various mechanisms and enumerate some of their common
characteristics. We can also draw useful conclusions
about the various categories and see how they compare
to each other.

@@

\begin{enumerate}
\item We provide an exploitation model that puts application attacks
  based on executable source code injection under a common
  perspective.
\item We categorize and analyze previously proposed defenses based on
  criteria that are considered critical when building protection
  mechanisms.
\item A worrying observation is that sometimes researchers test the
  diagnostic performance of their mechanisms in a poor manner. This
  turns out as very critical because some of these mechanisms can
  actually be bypassed.
\item Also worrying is that researchers do not make their mechanisms
  available to the community, thus no one can either use their
  mechanism, or replicate their experiments.
\item We discuss how our observations may be the main reasons why the
  mechanisms are not adopted in practice.
\item We discuss what are the necessary criteria for new mechanisms
  and how our work can help researchers to design new countermeasures.
\end{enumerate}

@@

Bratus et al.~\cite{BLSPS11} portray the issue in a generic fashion:
{\it ``unexpected (and unexpectedly powerful) computational models
  inside targeted systems, which turn a part of the target into a
  so-called `weird machine' programmable by the attacker via crafted
  inputs (a.k.a. `exploits')''}. In particular, {\it ``every
  application that copies untrusted input verbatim into an output
  program is vulnerable to code injection''}. Ray and
Ligatti~\cite{RL12b} actually proved the above claim based on formal
language theory.

@@

{\sc xss} attacks fall ino two categories, {\it non-persistent} and
{\it persistent}. Non-persistent {\sc xss} attacks take place when the
data provided by a user is used immediately by server-side scripts to
parse and display a page of results, without properly sanitizing the
request ({\sc step}: {\sc p-xss} 8 \text{\textbar} {\sc n-xss} 3).
Thus, in a non-persistent {\sc xss} attack the injected code
is never saved on the server and immediately becomes part of the
content that is sent back to a web user.
